\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-16 arch.tex
Added/refined material around the notion of planning and dispatch window for a reactor

This {\em synchronization} process propagates the current state
information at the execution frontier $\tau$ throughout all the
reactors. The sequencing of the reactors synchronization is identified
by traversing the reactor graph from the least dependents reactors --
the ones with no {\em external} timelines -- to the most dependent
ones which typicallu do not provide {\em internal} timelines.
Conversely each reactor can request a future change on one of its {\em
  external} state variables which will be transfromed as an {\em
  internal} goal to the owner.  Such \emph{goal dispatching} is
expected to produce a deliberation phase for the receiving reactor
which can cyclically and in turn, produce a plan impacting its {\em
  external} state variables' future that will propagate down following
the same mechanism.

@@

The number of flaws between two steps \comment{between synch. and
  execution?} can evolve for the following reasons that are exogenous
to deliberation:

\item A new goal has been posted on a reactor timeline. This is
  usually due to another reactor finding a partial plan and posting
  the corresponding new objective.

\item The outcome of synchronization created new flaws that did not
  need to be resolved during synchronization.

\item A previously requested goal has be recalled by the initial
  requester. This can occur when the reactor's initial plan is no
  longer valid, for instance with synchronization resulting in
  inconsistency.

\comment{this might be a subsubsection before someplace since it
  introduces the concept of latency/lookahead. TODO: yep it needs to
  move up to the architecture section.}
These events could potentially produce new flaws to be resolved while
any single deliberation step will likely reduce the number of flaws
for resolution. The scope of planning for a reactor also plays a role
in the deliberation process as we limit deliberation to the sliding
temporal window $[\tau+\lambda, \tau+\lambda+\pi]$ \cite{py10} where:

\item $\tau$ is the current tick indicative of the execution frontier

\item $\lambda$ is the specified latency of this reactor, which is an
  indicator of the maximum expected time for a reactor to produce a
  plan.\footnote{Note that while it is highly recommended to select a
    sound value, a failure to produce a plan within a chosen value for
    this parameter is not considered as a critical failure of the
    reactor.}

\item $\pi$ is the planning look-ahead of the reactor. which indicates
  how many ticks in the future this reactor is looking ahead during
  deliberation.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.65\columnwidth]{figs/plan_window}
  \caption{The planning window of the deliberation solver and its
    relation with the reactor latency and look-ahead. The dark grey tokens
    with dashed borders are excluded from deliberation as they do not
    overlap the interval $[\tau, \tau+\lambda+\pi]$}
  \label{fig:plan:window}
\end{figure}

As Fig. \ref{fig:plan:window} illustrates,  such a
scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable''
future. This window is taken into account by the \rx agent which will
notify a reactor of new {\em internal} goals not only when they
overlap this reactor specific window but also for the deliberation
solver in order to reduce the number of tokens to be evaluated to
reduce deliberation cost and performance. As time advances flaws that
were initially beyond this temporal horizon will become active which a
reactor can evaluate. This results in an apparent \emph{continuous
  planning} of the reactor. When there are no more flaws present in
the current planning scope a partial plan solution is found. This has
two effects:

>>>>

As the execution frontier -- called $\tau$ -- advance, the \rx agent
propagates the current state information accross all the reactors states
variables through the {\em synchronization} process.
 The sequencing of the reactors synchronization is identified
by traversing the reactor graph from the least dependents reactors --
the ones with no {\em external} timelines -- to the most dependent
ones which typicallu do not provide {\em internal} timelines.

Conversely each reactor can request a future change on one of its {\em
  external} state variables which will be transformed as an {\em
  internal} goal to the owner.  Such \emph{goal dispatching} is
expected to produce a deliberation phase for the receiving reactor
which can cyclically and in turn, produce a plan impacting its {\em
  external} state variables' future that will propagate down following
the same mechanism.

In order to avoid to be overflood by goals that may threats the
ability of a reactor to process these goals and integrate them in tis
plan. In order to do so each reactor specifies the two following
values that apply for the full scope of the agent lifetime:
\begin{itemize}
\item $\lambda$ is the specified latency of this reactor, which is an
  indicator of the maximum expected time for a reactor to produce a
  plan.\footnote{Note that while it is highly recommended to select a
    sound value, a failure to produce a plan within a chosen value for
    this parameter is not considered as a critical failure of the
    reactor.}
\item $\pi$ is the planning look-ahead of the reactor. which indicates
  how many ticks in the future this reactor is looking ahead during
  deliberation. 
\end{itemize}

Using this information along with the dependency relation betwen
reactor \rx defines the notion of execution latency of a state
variable as follow:
\begin{definition}
  The {\em execution latency}  of a {\em state variable} $s$ which is
    internal to the reactor $r$ is the maximum expected delay before 
    a goal on this state varabale can be included in its plan and
    executed by the  reactor $r$.

    It is defined recursrevely as follow: 
    \begin{equation*}
      \lambda_{exec}(s) = \lambda(r) + \max_{e \in exts(r)}
      \lambda_{exec}(e) 
  \end{equation*}
    Where  $exts(r)$ is the set of the state varaibles which are
    {\em externals} to $r$. 
\end{definition}

The concept of execution latency is critical at the agent level
and can also impact the planning/deliberation process on the {\em
  external} state variables of a reactor. For simplicity sake in the
future sections more focused in the \eu reactor, we abstract it out 
as being equivalent to the reactor latency (meaning that we 
assume that all {\em external} timelines of this reactor have
$\lambda_{exec} = 0$). At the \rx agent level though this
 execution latency is used in turn to define the {\em dispatch
window} of the same state variable:
\comment{Needed to say that early on but the flow of this paragraph 
is not very tight}
\begin{definition}
  \label{def:dispatch}
  For every timelines $s$ {\em internals} to the reactor $r$, the {\em 
    dispatch window} of this timeline at the execution frontier $\tau$
  is the time interval:
  \begin{equation*}
    \Pi(s) = [\tau + \lambda_{exec}(s), \tau+\lambda_{exec}(s)+\pi(r)]
  \end{equation*}
\end{definition}

This window is used by \rx in order to identify when to dispatch a
request to a {\em reactor} as a new goal. This happens only if the
request start time -- represented as the interval of possible start
instants -- intersects this window. Requests further in the future 
are buffered by the agent until this window advanced enough to 
overlap their start time. This gives a simple mechanism in order to
request a change from the {\em external} state variable of a a reactor 
while avoiding to dispatch too much goals at once to the reactor that
declare this same state variable as {\em internal}. The choice of both
the latency and look-ahead of a reactor can be delicate as one is
dependent on the other. Latency also highly dependent on exogenous
paramaters such as the computing power of the host system and the
number of reactors needing to deliberate at once. As of today these
should be evaluated carefully by the system designers and while
having an important impact in the efficacy of the agent to find and
execute a global strategy they are just indicative and a failure from
a reactor to produce a plan in its latency is not considered as a
critical failure from \rx point of view.

Requests from a reactor can  be {\em recalled} by this reactor when it
they are not part of its plan anymore. For example if this reactor
need to replan due to an unexpected event, it will recall ol the
requests it did propduce that have not been observed yet. These
recalls are notified to the reactor that received this request as a
goal in order to remove it from its current objectives.

@@

{\em Synchronization} and the dispatching of new {\em goals} from
\rx can occur between two steps and alter the plan. These alterations
exogenous to the deliberation process are dynamically impacting the set
of flaws that need to be resolved. Obviously new flaws are created
when a new goal is received from \rx and conversely often the recall
of an objective will remove the corresponding flaws\footnote{although
  if the flaws of this gaol were previously resolved, the removal of
  this goal may create new flaws}. The impact of {\em synchronization}
can be more subtle as its alteration of the plan at the execution
frontier is much more drastic. We can still give a non exhaustive list
of possible alterations of the flaws directly related to {\em
  synchronization}:
\begin{enumerate}
\item Previous flaws directly related to the execution frontier have
  been resolved by the {\em synchronization} solver.
\item The {\em synchronization} solver may have created new flaws that
  were outside of its scope but are now part of the {\em deliberation}
  solver scope.
\item Last observations proved the current plan to be invalid
  requiring the {\em synchronization} solver to relax the existing
  plan in order to identify its current {\em external} state.
\end{enumerate}
Finally as {\em synchronization} occurs when the {\em
  execution frontier} advances, it implies that the reactor planning
scope is further in the future. This need to be reflected by the {\em
  deliberation} solver which need a new deliberation scope one tick
further in the future.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.65\columnwidth]{figs/plan_window}
  \caption{\small The planning window of the deliberation solver and its
    relation with the {\em dispatch window}. The dark grey
    tokens   with dashed borders are excluded from deliberation as they do not
    overlap the interval $[\tau, \tau+\lambda+\pi]$. All the other
    tokens (in {\color{green} green}) -- along with the goals received from the
    agent (in {\color{blue} blue}) which can overlap this window are
    evaluated.}
  \label{fig:plan:window}
\end{figure}

As Fig. \ref{fig:plan:window} illustrates,  such a
scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable''
future. This filtering window complements the \rx {\em dispatch
  window} from definition \ref{def:dispatch} which already filters
out at the agent level the goals that are before the reactor execution
latency or too far in the future. The planning window inside of the
reactor refines further this filtering as tokens can be excluded
dynamically from deliberation as their insertion in the plan and
scheduling with other tokens generated can reduced their temporal
scope outside of the planning scope. This helps the deliberation 
solver reduce the number of tokens to be evaluated and therefore
reduce deliberation cost and performance. As time advances flaws that
were initially beyond this temporal horizon will become active which a
reactor can evaluate. This results in an apparent \emph{continuous
  planning} of the reactor. When there are no more flaws present in
the current planning scope a partial plan solution is found. This has
two effects:
